Contribution Title\thanks{Supported by organization x.}

### 2019-10-20 paper/paper.tex
EOD

\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
%% \usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%%                         %% http://ctan.org/pkg/subcaption
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{tikz-qtree} % Used for syntax trees
\usepackage{tikz-cd} % Used for commutative diagrams

@@

\noindent The absence of both non-grouping parentheses and marks mean that every permissible occurrence of a pair of parentheses must permit zero or more pairs. Adding marks adds required pairs for some trees, turning each occurrence to one of $k$ or more pairs, for some non-negative integer $k$ determined by the tree. The left column is more complicated since non-grouping parentheses can introduce occurrences that require an exact number of parentheses, as well as all the complications of marks.

@@

\begin{description}
  \item[$w_\top$ is ambiguous.] In this case, $\exists t'.\ t \neq t' \land t' \in \parse(w_\top)$. This means that we can take the canonical encoding of $t'$ (call it $c'$) and replace every ``$[]$'' with either one or zero parentheses and produce $w_\top$. But that also means that every word we construct from $c$ (by choosing some non-negative integer of parentheses for each ``$[]$'') we can construct the same word from $c'$ (by choosing the same number for corresponding parentheses, and zero for the others), i.e., $\words(t) \subseteq \words(t')$.
  \item[$w_\top$ is unambiguous.] In this case $\words(t)$ has at least one unambiguous word ($w_\top$) which thus cannot be part of $\words(t')$ for some other $t' \in L(T_D)$, thus $\neg \exists t'.\ t \neq t' \land \words(t) \subseteq \words(t')$.
\end{description}

@@

Adding marks to such a language preserves resolvability:

\begin{itemize}
  \item Marks can introduce at most one required pair of parentheses per node.
  \item Adding marks to a language can only ever shrink $\words(t)$ for any particular $t \in L(T_D)$.
  \item Given a tree $t$, the word produced by putting one pair of parentheses around each node in $t$ (potentially adding double parentheses), call it $w$, is unambiguous in the language without marks. It cannot be excluded from $\words(t)$ by adding marks (since it already has a pair of parentheses around each node), and it also cannot be ambiguous, since that would require $\words(t')$ for some other tree $t'$ to have grown. Thus, $\parse(w) = \{t\}$ even when we add marks to $D$.
\end{itemize}

@@

\begin{center}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,
    scale = 1,transform shape]
  \node[state] (1) {$1$};
  \node[state] (2) [right of=1] {$2$};

  \node[state] (3) [below of=1,xshift=-1.2cm] {$3$};
  \node[state] (4) [right of=3] {$4$};
  \node[state] (5) [right of=4] {$5$};

  \node[state] (6) [below of=1,yshift=-2cm] {$6$};
  \node[state] (7) [right of=6] {$7$};

  \node[state] (8) [below of=6] {$8$};
  \node[state] (9) [right of=8] {$9$};

  \node[state,initial] (s) [left of=1,xshift=-1cm] {$s$};
  \node[state,accepting] (f) [right of=2,xshift=1cm] {$f$};

  \path (1) edge node[left]{$'\pospl', +(1, 2)$} (3)
  (5) edge[right] node{$'\pospr', -(1, 2)$} (2);
  \path (1) edge[out=180,in=180,looseness=2,left] node{$'\pospl', +(1, 2)$} (6)
  (7) edge[out=0,in=0,looseness=2,right] node{$'\pospr', -(1, 2)$} (2);

  \path (3) edge node{$'+'$} (4);
  \path (4) edge[bend right,above,out=270,in=225] node[xshift=5mm]{$'\pospl', +(4, 5)$} (3)
  (5) edge[loop right,below] node[yshift=-3.5mm]{$'\pospr', -(4, 5)$} (5);
  \path (4) edge[left] node{$'\pospl', +(4, 5)$} (6)
  (7) edge node{$'\pospr', -(4, 5)$} (5);

  \path (6) edge[left] node{$'\pospl', +(6, 7)$} (8)
  (9) edge[right] node{$'\pospr', -(6, 7)$} (7);

  \path (8) edge[above] node{$'n'$} (9);

  \path (s) edge[bend left] node{$'\pospl', +(s, f)$} (1)
  (2) edge[bend left] node{$'\pospr', -(s, f)$} (f);

\end{tikzpicture}
\end{center}

\noindent We then combine these disparate DFAs to produce a single visibly pushdown automaton in three steps:

@@

\noindent To make the figure less cluttered we omit $'\pospl'$ and $'\pospr'$, i.e., we abbreviate $p \xrightarrow{'\pospl', +(p',q')} q$ as $p \xrightarrow{+(p', q')}$ and $p \xrightarrow{'\pospr', -(p',q')} q$ as $p \xrightarrow{-(p', q')}$.

\begin{center}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,
    scale = 1,transform shape]
  \node[state] (1) {$1$};
  \node[state] (2) [right of=1] {$2$};

  \node[state] (3) [below of=1,xshift=-1.2cm] {$3$};
  \node[state] (4) [right of=3] {$4$};
  \node[state] (5) [right of=4] {$5$};

  \node[state] (6) [below of=1,yshift=-2cm] {$6$};
  \node[state] (7) [right of=6] {$7$};

  \node[state] (8) [below of=6] {$8$};
  \node[state] (9) [right of=8] {$9$};

  \node[state,initial] (s) [left of=1,xshift=-1cm] {$s$};
  \node[state,accepting] (f) [right of=2,xshift=1cm] {$f$};

  \path (1) edge node[left]{$'\pospl, +(1, 2)$} (3)
  (5) edge[right] node{$'\pospr', -(1, 2)$} (2);
  \path (1) edge[out=180,in=180,looseness=2,left] node{$'\pospl', +(1, 2)$} (6)
  (7) edge[out=0,in=0,looseness=2,right] node{$'\pospr', -(1, 2)$} (2);

  \path (3) edge node{$'+'$} (4);
  \path (4) edge[bend right,above,out=270,in=225] node[xshift=5mm]{$'\pospl', +(4, 5)$} (3)
  (5) edge[loop right,below] node[yshift=-3.5mm]{$'\pospr', -(4, 5)$} (5);
  \path (4) edge[left] node{$'\pospl', +(4, 5)$} (6)
  (7) edge node{$'\pospr', -(4, 5)$} (5);

  \path (6) edge[left] node{$'\pospl', +(6, 7)$} (8)
  (9) edge[right] node{$'\pospr', -(6, 7)$} (7);

  \path (8) edge[above] node{$'n'$} (9);

  \path (s) edge[bend left] node{$'\pospl', +(s, f)$} (1)
  (2) edge[bend left] node{$'\pospr', -(s, f)$} (f);

\end{tikzpicture}
\end{center}

@@

\begin{center}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,
    scale = 1,transform shape]
  \node[state] (1) {$1$};
  \node[state] (2) [right of=1] {$2$};

  \node[state] (3) [below of=1,xshift=-1.2cm] {$3$};
  \node[state] (4) [right of=3] {$4$};
  \node[state] (5) [right of=4] {$5$};

  \node[state] (6) [below of=1,yshift=-2cm] {$6$};
  \node[state] (7) [right of=6] {$7$};

  \node[state] (8) [below of=6] {$8$};
  \node[state] (9) [right of=8] {$9$};

  \node[state,initial] (s) [left of=1,xshift=-1cm] {$s$};
  \node[state,accepting] (f) [right of=2,xshift=1cm] {$f$};

  \path (1) edge node[left]{$'\pospl, +(1, 2)$} (3)
  (5) edge[right] node{$'\pospr', -(1, 2)$} (2);
  \path (1) edge[out=180,in=180,looseness=2,left] node{$'\pospl', +(1, 2)$} (6)
  (7) edge[out=0,in=0,looseness=2,right] node{$'\pospr', -(1, 2)$} (2);

  \path (3) edge node{$'+'$} (4);
  \path (4) edge[bend right,above,out=270,in=225] node[xshift=5mm]{$'\pospl', +(4, 5)$} (3)
  (5) edge[loop right,below] node[yshift=-3.5mm]{$'\pospr', -(4, 5)$} (5);
  \path (4) edge[left] node{$'\pospl', +(4, 5)$} (6)
  (7) edge node{$'\pospr', -(4, 5)$} (5);

  \path (6) edge[left] node{$'\pospl', +(6, 7)$} (8)
  (9) edge[right] node{$'\pospr', -(6, 7)$} (7);

  \path (8) edge[above] node{$'n'$} (9);

  \path (s) edge[bend left] node{$'\pospl', +(s, f)$} (1)
  (2) edge[bend left] node{$'\pospr', -(s, f)$} (f);

\end{tikzpicture}
\end{center}

@@

    \delta(p, \text{``}\pospl\text{''}, \lambda)
    & \{ (s_{l_3}, (p, q, w))
    & p \in Q_{l_1}, \delta_{l_1}(p, l_2) = q, (l_2, l_3, w) \in T \} \\

    \delta(s, \text{``}\pospl\text{''}, \lambda)
    & \{ (s_{l_2}, (s, f, w))
    & \labelnt(l_1) = S, (l_1, l_2, w) \in T \} \\
  \end{array}
  $$
\end{itemize}

@@

\noindent To find two distinct successful runs, one in each automaton, that recognize the same word we construct the product automaton $A_{\posp{}} \times A'_{\posp{}} = (Q \times Q, \T', \Gamma \times (\Gamma \cup \{\gamma\}), \delta_{\times}, (s, s), \{(f, f)\})$, where $\delta_{\times}$ is given in Section~\ref{sec:preliminaries-vpls} (using the partitions $\T_c = \{ \pospl \}$, $\T_i = \T$, and $\T_r = \{\pospr\}$). If the product automaton has a successful run that passes through at least one configuration $((p, p'), (g, g')\cdot w)$ such that $p \neq p' \lor g \neq g'$ (distinct states, or distinct stack symbols, respectively), then that run corresponds to two distinct runs in $A_{\posp{}}$ and $A'_{\posp{}}$. We check for the existence of such a run by trimming the product automaton and looking for a transition that pushes $(g, g')$ where $g \neq g'$, or transitions to $(q, q')$ where $q \neq q'$.

Performing this procedure directly on a language definition $D$ gives a sound decision procedure where the presence of two distinct runs implies unresolvably ambiguous. If we instead do it on a modified $D$ with no marks on the right hand side of any production (i.e., all non-terminals have the form $N_\emptyset$) we get a sound decision procedure where the absence of two distinct runs implies resolvably ambiguous.

There are thus cases where we can answer resolvably ambiguous or unresolvably ambiguous with certainty, and cases where we cannot answer either way with certainty.

@@

\begin{lemma}
  All words $w \in L(A_{\posp{}})$ are well-formed linear encodings.
\end{lemma}

@@

\begin{lemma}
  $\forall t \in L(T_D).\ \lattice(\runtoword(\treetorun(t))) = \lattice(t)$.
\end{lemma}

\begin{lemma}
  For all $t \in L(T_D)$ and every well-formed lattice encoding $w$ such that $\lattice(t) \subseteq \lattice(w)$, $w \in L(A'_{\posp{}})$.
\end{lemma}

\begin{lemma}
  We can define a function $\runprimetorun$ from runs in $A'_{\posp{}}$ to runs in $A_{\posp{}}$ such that $\forall R.\ \lattice(\runtoword(\runprimetorun(R))) \subseteq \lattice(\runtoword(R))$.
\end{lemma}

\begin{proof}
  By removing every transition that pushes or pops $\gamma$ and replacing every used transition $p \xrightarrow{'\pospl', +g} q$ that does not exist in $A_{\posp{}}$ with a transition $p \xrightarrow{'\reqpl', +g} q$.
\end{proof}

\begin{theorem}
  The existence of two distinct runs $R$ and $R'$, where $R$ is a successful run in $A_{\posp{}}$ and $R'$ in $A'_{\posp{}}$, and $\runtoword(R) = \runtoword(R')$, implies the existence of two distinct trees such that the lattice of one is entirely contained in the other.
\end{theorem}

@@

  Since $\runprimetorun(R') \neq R$ we have two distinct trees $t_1 = \runtotree(\runprimetorun(R'))$ and $t_2 = \runtotree(R)$. Since $\lattice(\runtoword(\runprimetorun(R'))) \subseteq \lattice(\runtoword(R'))$ and $\lattice(\runtoword(R')) = \lattice(\runtoword(R))$ we have that $\lattice(t_1) \subseteq \lattice(t_2)$.
\end{proof}

@@

The dynamic resolvability problem is as follows: for a given $w' \in L(G'_D)$ determine whether $\forall t \in \parse(w').\ \exists w'_2.\ \parse(w'_2) = \{t\}$. Furthermore, for practical reasons, if the word is resolvably ambiguous we wish to produce a (minimal) witness for each tree. Similarly to Section~\ref{sec:static}, we place some restrictions on $D$, though the restrictions are less severe. While Section~\ref{sec:static} requires a language definition $D$ to not contain parentheses, this section merely requires parentheses to be balanced. Concretely, the restrictions are as follows:

@@

\noindent Both of these restrictions can be checked statically.

>>>>

\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
%% \usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%%                         %% http://ctan.org/pkg/subcaption
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,automata,calc}
\usepackage{tikz-qtree} % Used for syntax trees
\usepackage{tikz-cd} % Used for commutative diagrams

@@

\noindent The absence of both non-grouping parentheses and marks mean that every occurrence of a pair of parentheses must permit zero or more pairs. Adding marks adds required pairs for some trees, turning each occurrence to one of $k$ or more pairs, for some non-negative integer $k$ determined by the tree. The left column is more complicated since non-grouping parentheses can introduce occurrences that require an exact number of parentheses, as well as all the complications of marks.

@@

\begin{description}
  \item[$w_\top$ is ambiguous.] In this case, $\exists t'.\ t \neq t' \land t' \in \parse(w_\top)$. This means that we can take the canonical encoding of $t'$ (call it $c'$) and replace every ``$[]$'' with either one or zero parentheses and produce $w_\top$. But that also means that every word we construct from $c$ (by choosing some non-negative integer of parentheses for each ``$[]$'') can also be constructed from $c'$ (by choosing the same number for corresponding parentheses, and zero for the others), i.e., $\words(t) \subseteq \words(t')$.
  \item[$w_\top$ is unambiguous.] In this case $\words(t)$ has at least one unambiguous word ($w_\top$) which thus cannot be part of $\words(t')$ for any other $t' \in L(T_D)$, thus $\neg \exists t'.\ t \neq t' \land \words(t) \subseteq \words(t')$.
\end{description}

@@

Adding marks to such a language preserves resolvability, since:

\begin{itemize}
  \item Marks can introduce at most one required pair of parentheses per node.
  \item Adding marks to a language can only ever shrink $\words(t)$ for any particular $t \in L(T_D)$.
  \item Given a tree $t$, the word produced by putting one pair of parentheses around each node in $t$ (thus potentially adding double parentheses), call it $w$, is unambiguous in the language without marks. It cannot be excluded from $\words(t)$ by adding marks (since it already has a pair of parentheses around each node), and it also cannot be ambiguous, since that would require $\words(t')$ for some other tree $t'$ to have grown. Thus, $\parse(w) = \{t\}$ even when we add marks to $D$.
\end{itemize}

@@

\begin{center}
\begin{tikzpicture}[->,>=Latex,shorten >=1pt,auto,node distance=2cm,
    scale = 1,transform shape]
  \node[state,initial] (1) {$1$};
  \node[state,accepting] (2) [right of=1] {$2$};
  \node[state,accepting] (3) [right of=2] {$3$};

  \node[state,initial] (4) [below of=1,yshift=7mm] {$4$};
  \node[state] (5) [right of=4] {$5$};
  \node[state,accepting] (6) [right of=5] {$6$};

  \node[state,initial] (7) [below of=4,yshift=7mm] {$7$};
  \node[state,accepting] (8) [right of=7] {$8$};

  \path (1) edge node{$E$} (2)
  (2) edge node{$';'$} (3);

  \path (4) edge node{$E$} (5)
  (5) edge node{$'s'$} (6);

  \path (7) edge node{$'z'$} (8);
\end{tikzpicture}
\end{center}

\noindent We then combine these separate DFAs to produce a single visibly pushdown automaton in three steps:

@@

\noindent To make the figure less cluttered we omit $'\pospl'$ and $'\pospr'$, i.e., we abbreviate $p \xrightarrow{'\pospl', +(p',q')} q$ as $p \xrightarrow{+(p', q')} q$ and $p \xrightarrow{'\pospr', -(p',q')} q$ as $p \xrightarrow{-(p', q')} q$.

\begin{center}
\begin{tikzpicture}[->,>=Latex,shorten >=1pt,auto,node distance=2cm,
    scale = 1,transform shape]
  \node[state] (1) {$1$};
  \node[state] (2) [right of=1] {$2$};
  \node[state] (3) [right of=2] {$3$};

  \node[state] (4) [below of=1] {$4$};
  \node[state] (5) [right of=4] {$5$};
  \node[state] (6) [right of=5] {$6$};

  \node[state] (7) [below of=4] {$7$};
  \node[state] (8) [right of=7] {$8$};

  \node[state,initial] (s) [left of=1,xshift=-0.9cm] {$s$};
  \node[state,accepting] (f) [right of=3,xshift=0.9cm] {$f$};

  \path (2) edge node{$';'$} (3);

  \path (5) edge node{$'s'$} (6);

  \path (7) edge node{$'z'$} (8);

  \path (s) edge node{$+(s, f)$} (1);
  \draw (2.north) -- ([yshift=4mm]2.north) -| (f.north);
  \draw (3.north) -- ([yshift=4mm]3.north) -| (f.north);
  \node at ([yshift=5mm]$(3)!0.5!(f)$) {$-(s,f)$};

  % 1 -E-> 2
  \draw (1.east) -| ([xshift=4mm]4.east) |- ([xshift=4mm,yshift=10mm]7.east) -- (7);
  \draw ([xshift=4mm]4.east) -- (4);
  \draw (6.north) |- ([xshift=-8mm]$(2)!0.5!(5)$) |- (2.west);
  \draw (8) -- ([xshift=-3.5mm,yshift=10mm]8.west) |- (2.west);
  \node at ([xshift=2mm,yshift=2mm]$(1)!0.5!(4)$) {$+(1,2)$};
  \node at ([xshift=-2mm,yshift=2mm]$(2)!0.5!(5)$) {$-(1,2)$};

  % 4 -E-> 5
  \draw (4.south) -- (7.north);
  \path (4) edge[out=270,in=225,looseness=5] node[xshift=4mm]{$+(4,5)$} (4);
  \draw (6.south) -- ([yshift=-5mm]6.south) -| (5.south);
  \path (8) edge[right] node[yshift=-3mm]{$-(4,5)$} (5);
\end{tikzpicture}
\end{center}

@@

\begin{center}
\begin{tikzpicture}[->,>=Latex,shorten >=1pt,auto,node distance=2cm,
    scale = 1,transform shape]
  \node[state] (1) {$1$};
  \node[state] (2) [right of=1] {$2$};
  \node[state] (3) [right of=2] {$3$};

  \node[state] (4) [below of=1] {$4$};
  \node[state] (5) [right of=4] {$5$};
  \node[state] (6) [right of=5] {$6$};

  \node[state] (7) [below of=4] {$7$};
  \node[state] (8) [right of=7] {$8$};

  \node[state,initial] (s) [left of=1,xshift=-0.9cm] {$s$};
  \node[state,accepting] (f) [right of=3,xshift=0.9cm] {$f$};

  \path (2) edge node{$';'$} (3);

  \path (5) edge node{$'s'$} (6);

  \path (7) edge node{$'z'$} (8);

  \path (s) edge node{$+(s, f, \epsilon)$} (1)
  (3) edge node{$-(s, f, \epsilon)$} (f);

  % 1 -E-> 2
  \draw (1.east) -| ([xshift=4mm]4.east) |- ([xshift=4mm,yshift=10mm]7.east) -- (7);
  \draw ([xshift=4mm]4.east) -- (4);
  \draw (6.north) |- ([xshift=-8mm]$(2)!0.5!(5)$) |- (2.west);
  \draw (8) -- ([xshift=-3.5mm,yshift=10mm]8.west) |- (2.west);
  \node at ([xshift=1mm,yshift=2mm]$(1)!0.5!(4)$) {$+(1,2,\epsilon)$};
  \node at ([xshift=-1mm,yshift=2mm]$(2)!0.5!(5)$) {$-(1,2,\epsilon)$};

  % 4 -E-> 5
  \draw (4.south) -- (7.north);
  \path (4) edge[out=270,in=225,looseness=5] node[xshift=4mm]{$+(4,5,\epsilon)$} (4);
  \draw (6.south) -- ([yshift=-5mm]6.south) -| (5.south);
  \path (8) edge[right] node[yshift=-3mm]{$-(4,5,\epsilon)$} (5);

  % s -E(t)-> f
  \draw (s.south) |- (7.west);
  \draw (s.south) |- (4.west);
  \node at ([xshift=8mm,yshift=-10mm]s) {$+(s, f, a)$};
  \draw (8.east) -| (f.south);
  \draw (6.east) -| (f.south);
  \node at ([xshift=-8mm,yshift=-10mm]f) {$-(s, f, a)$};
\end{tikzpicture}
\end{center}

@@

    \delta(s, \text{``}\pospl\text{''}, \lambda)
    & \{ (s_{l}, (s, f, w))
    & (S, w, N) \in T, N = \labelnt(l) \} \\

    \delta(p, \text{``}\pospl\text{''}, \lambda)
    & \{ (s_{l_3}, (p, q, w))
    & p \in Q_{l_1}, \delta_{l_1}(p, N) = q, (N, w, N') \in T, N' = \labelnt(l_3) \} \\

    \delta(p, \text{``}\pospr\text{''}, (s, f, w))
    & \{ (f, \lambda)
    & (S, w, N) \in T, N = \labelnt(l), p \in F_l \} \\

    \delta(q', \text{``}\pospr\text{''}, (p, q, w))
    & \{ (q, \lambda)
    & q' \in F_{l_1}, \labelnt(l_1) = N', (N, w, N') \in T, p \in Q_{l_2}, \delta_{l_2}(p, N) = q \} \\

  \end{array}
  $$
\end{itemize}

@@

\noindent To find two distinct successful runs, one in each automaton, that recognize the same word we construct the product automaton $A_{\posp{}} \times A'_{\posp{}} = (Q \times Q, \T', \Gamma \times (\Gamma \cup \{\gamma\}), \delta_{\times}, (s, s), \{(f, f)\})$, where $\delta_{\times}$ is described in \cite{alurVisiblyPushdownLanguages2004} (using the partitions $\T_c = \{ \pospl \}$, $\T_i = \T$, and $\T_r = \{\pospr\}$). If the product automaton has a successful run that passes through at least one configuration $((p, p'), (g, g')\cdot w)$ such that $p \neq p' \lor g \neq g'$ (distinct states, or distinct stack symbols, respectively), then that run corresponds to two distinct successful runs in $A_{\posp{}}$ and $A'_{\posp{}}$. We check for the existence of such a run by trimming the product automaton (as described in \cite{caralpTrimmingVisiblyPushdown2015}) and looking for a transition that pushes $(g, g')$ where $g \neq g'$, or transitions to $(q, q')$ where $q \neq q'$.

Performing this procedure on a language definition $D$ without non-grouping parentheses and marks gives a sound and complete decision procedure for determining if $D$ is resolvably ambiguous. If we instead take a language definition $D$ with marks but no non-grouping parentheses, then remove the marks and perform the procedure, then the absence of two such runs implies that $D$ is resolvably ambiguous, while their presence only implies the possibility of an unresolvable ambiguity; the marks might remove it.

@@

@@

\begin{lemma}
  $\forall t \in L(T_D).\ \runtoword(\treetorun(t))$ is the canonical linear encoding of $\words(t)$.
\end{lemma}

\begin{lemma}
  For all $t \in L(T_D)$ and every linear encoding $c$ such that $\words(t) \subseteq L(c)$, $c \in L(A'_{\posp{}})$.
\end{lemma}

\begin{lemma}
  We can define a function $\runprimetorun$ from runs in $A'_{\posp{}}$ to runs in $A_{\posp{}}$ such that $\forall R.\ L(\runtoword(\runprimetorun(R))) \subseteq L(\runtoword(R))$.
\end{lemma}

\begin{proof}
  By removing every transition that pushes or pops $\gamma$.
\end{proof}

\begin{theorem}
  The existence of two distinct runs $R$ and $R'$ such that $R$ is a successful run in $A_{\posp{}}$ and $R'$ in $A'_{\posp{}}$, and $\runtoword(R) = \runtoword(R')$, implies the existence of two distinct trees such that the word language of one is entirely contained in the other.
\end{theorem}

@@

  Since $\runprimetorun(R') \neq R$ we have two distinct trees $t_1 = \runtotree(\runprimetorun(R'))$ and $t_2 = \runtotree(R)$. Since $L(\runtoword(\runprimetorun(R'))) \subseteq L(\runtoword(R'))$ and $L(\runtoword(R')) = L(\runtoword(R))$ we have that $L(t_1) \subseteq L(t_2)$.
\end{proof}

We thus have a way to detect the presence or absence of a pair of trees $t_1, t_2 \in L(T_D)$ such that $\words(t_1) \subseteq \words(t_2)$, which completely determines the resolvability of $D$ when $D$ has no non-grouping parentheses or marks.

@@

The dynamic resolvability problem is as follows: for a given $w' \in L(G'_D)$ determine whether $\forall t \in \parse(w').\ \exists w'_2.\ \parse(w'_2) = \{t\}$. Furthermore, for practical reasons, if the word is resolvably ambiguous we wish to produce a (minimal) witness for each tree. Similarly to Section~\ref{sec:static}, we place some restrictions on $D$, though the restrictions are less severe. While Section~\ref{sec:static} requires $D$ to have no non-grouping parentheses and no marks, this section has two much less stringent requirements:

@@

\noindent Both of these restrictions can be checked statically. Additionally, if we already know statically that $D$ is resolvably ambiguous\footnote{While the \emph{decision} part of this procedure is uninteresting in this case (it will always answer ``resolvable''), the minimal witness \emph{is} interesting: it shows the user how to resolve the ambiguity.} we can drop the second requirement.

Similarly to Section~\ref{sec:static} the approach centers around the word language $\words(t)$. We can easily construct a visibly pushdown automaton for this language by going via the canonical linear encoding: let every pair ``$\reqp{}$'' push and pop the same stack symbol (call it $\gamma$) and let every ``$\posp{}$'' push and pop a unique stack symbol. For example, ``$\posp{\reqp{a}\posp{b}}$'' produces the following automaton:

\begin{center}
\begin{tikzpicture}[->,>=Latex,shorten >=1pt,auto,node distance=2cm,
    scale = 1,transform shape]
  \node[state,initial] (1) {};
  \node[state] (2) [right of=1] {};
  \node[state] (3) [right of=2] {};
  \node[state] (4) [right of=3] {};
  \node[state,accepting] (5) [right of=4] {};

  \path (1) edge[loop above] node{$'(', +1$} (1)
  (1) edge node{$'(', +\gamma$} (2)
  (2) edge node{$'a'$} (3)
  (3) edge node{$')', -\gamma$} (4)
  (4) edge[loop above] node{$'(', +2$} (4)
  (4) edge node{$'b'$} (5)
  (5) edge[loop above] node{$')', -2$} (5)
  (5) edge[loop right] node{$')', -1$} (5);
\end{tikzpicture}
\end{center}

\noindent This is useful to us since visibly pushdown automata are closed under difference. Given a pair of trees $t_1$ and $t_2$ we can thus produce a pair of automata recognizing $\words(t_1)$ and $\words(t_2)$ respectively, then construct a pair of automata recognizing $\words(t_1) \setminus \words(t_2)$ and $\words(t_2) \setminus \words(t_1)$, respectively. These new automata recognize the words that are not ambiguous between these two particular trees. Our algorithm thus takes the following shape:

\begin{verbatim}
Proc(T)
  let R = \emptyset
  let U = \emptyset
  for t \in T
    let T' = T \setminus \{t\}
    let A = \words(t)
    loop
      A = A \setminus \bigcup_{t' \in T'} \words(t')
      if L(A) = \emptyset
        U = U \cup \{t\}
        break loop
      if \parse(\shortest(A)) = \{t\}
        R = R \cup \{(t, \shortest(A))\}
        break loop
      T' = \parse(\shortest(A)) \setminus \{t\}
  return (R, U)
\end{verbatim}

%% TODO: CONTINUE
